---
title: "Bayesian linear regression"
author: "Marni Tausen"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.format='html', comment='')
set.seed(600)
```

### blm package

The blm package is an R package for Bayesian linear models, which has
the same kind of interface as lm and glm in R. The difference of the
Bayesian model, is that we do not directly estimate the value of the
parameters, instead we estimate an normal distribution giving the
mean and the variance for each of the parameters.

### Installation

The package is currently only available through github. Therefore the
easiest way to install the package, is to use the devtools package.
Skip the installation part, if already installed.
```{r, eval=FALSE}
install.packages("devtools")
library(devtools)
```

Using devtools, you can install the package using install_github().

```{r, eval=FALSE}
install_github("MarniTausen/blm")
```

and voila! The package is installed.

### Quick Guide

The blm package uses the same interface as the lm and glm functions in
R, and therefore functions very similarly.

First to load the package:

```{r}

library(blm)

```

```{r, echo=FALSE}
x <- 1:100
y <- rnorm(100, x)
```

To construct the blm model use the blm function and feed it a
formula, with the structure **response** ~ **predictors**.
```{r}

blm(y ~ x)

```
To extract summary statistics of the blm model, we can combine
```{r}

summary(blm(y ~ x))

```

A quick guide of how to use the package.
Short and simple. Very straight forward.

### Features

A list of features the package contains, and a short description of
their purpose and function. For a more indepth description check the
documentation with ?x.blm, or the github wiki page for the function.

* blm
	* Description
* coef (coefficients)
	* Description
* confint
	* Description
* deviance
	* Description
* fitted
	* Description
* predict
	* Description
* residuals
	* Description
* summary
	* Description
